### 2021-08-31 scripts/R/plot_results.rmd
add GWAS TPR simulation code, runnable in docker

---
title: "Plot results"
author: nadeaus
date: 31.08.21
output: html_notebook
---

Setup
```{r}
require(dplyr)
require(ggplot2)
source("scripts/R/functions/plot_functions.R")
accuracy_simulation_results <- read.delim("output/2021-08-31_estimator_accuracy_MLE.txt")

View(accuracy_simulation_results %>% filter(alpha == 5))
```

Simulation accuracy heatmap
```{r}
# Summarize results for plotting
error.cols <- c("RMSE.z.h", "RMSE.h.MWA.h")

data.long <- do.call(
  what = tidyr::gather,
  args = c(error.cols, list(data = accuracy_simulation_results, key = "inference.method", value = "error")))

data.summary <- as.data.frame(
  data.long %>%
    dplyr::group_by(inference.method, alpha, H2, var.z, p) %>%
    dplyr::summarise(mean.error = mean(error), sd.error = sd(error), N.reps = dplyr::n()))

# Create facet variable labels for plot
data.summary$inference.method <- factor(
  data.summary$inference.method,
  levels = c("RMSE.h.MWA.h", "RMSE.z.h"),
  labels = c("'Estimated env. part of spVL: '*bolditalic(hat(e))",
             "'Measured spVL, scaled: '*bolditalic(z) - bold(bar(z))"))

# Plot RMSE
rmse_plot <- plotHeatmap(
  df = data.summary,
  N.sim = data.summary$N.reps[[1]],
  is.MLE = T,
  scale.limits = c(0, 1),
  legend.name = "RMSE",
  midpoint = 0.5,
  fill.hi = RColorBrewer::brewer.pal(name = 'PiYG', n = 3)[1],
  fill.lo = RColorBrewer::brewer.pal(name = 'PiYG', n = 3)[3],
  fill.mid = RColorBrewer::brewer.pal(name = 'PiYG', n = 3)[2],
  x.var = "alpha",
  y.var = "H2",
  fill.var = "mean.error",
  facet.x = "inference.method",
  facet.y = NA,
  is.oob.squish = F,
  is.real.scale = F,
  labeller = label_parsed) +
  labs(x = expression("Selection strength (" * alpha * ")"),
       y = expression("spVL heritability (" * italic('H'^2) * ")"),
       title = "A") +
  theme(
    text = element_text(size = 8),
    panel.border = element_rect(colour = "black", size = 0.5, fill = NA),
    strip.background = element_rect(colour = "black", size = 0.5))

show(rmse_plot)

ggsave(plot = rmse_plot,
       filename = "manuscript/figures/simulation_RMSE.png",
       width = 4.25,
       height = 2,  # maintain height as 2 for consistent font size in manuscript
       units = "in")
```