### 2021-09-28 scripts/R/plot_results.rmd
migrate code to make full figure 1, add figures in manuscript directory

---
title: "Plot results"
author: nadeaus
date: 31.08.21
output: html_notebook
---

Setup
```{r}
knitr::opts_knit$set(root.dir = "/Users/nadeaus/Repos/poumm-gwas")
```

```{r}
require(dplyr)
require(ggplot2)
source("scripts/R/functions/plot_functions.R")
source("scripts/R/functions/utility_functions.R")
accuracy_simulation_results <- read.delim("output/2021-08-31_estimator_accuracy_MLE.txt")
tpr_simulation_results <- read.delim("output/2021-08-31_GWAS_TPR_MLE.txt")
```

Simulation accuracy heatmap
```{r}
# Summarize results for plotting
error.cols <- c("RMSE.z.h", "RMSE.h.MWA.h")

data.long <- do.call(
  what = tidyr::gather,
  args = c(error.cols, list(data = accuracy_simulation_results, key = "inference.method", value = "error")))

data.summary <- as.data.frame(
  data.long %>%
    dplyr::group_by(inference.method, alpha, H2, var.z, p) %>%
    dplyr::summarise(mean.error = mean(error), sd.error = sd(error), N.reps = dplyr::n()))

# Create facet variable labels for plot
data.summary$inference.method <- factor(
  data.summary$inference.method,
  levels = c("RMSE.h.MWA.h", "RMSE.z.h"),
  labels = c("'Estimated env. part of spVL: '*bolditalic(hat(e))",
             "'Measured spVL, scaled: '*bolditalic(z) - bold(bar(z))"))

# Plot RMSE
rmse_plot <- plotHeatmap(
  df = data.summary,
  N.sim = data.summary$N.reps[[1]],
  is.MLE = T,
  scale.limits = c(0, 1),
  legend.name = "RMSE",
  midpoint = 0.5,
  fill.hi = RColorBrewer::brewer.pal(name = 'PiYG', n = 3)[1],
  fill.lo = RColorBrewer::brewer.pal(name = 'PiYG', n = 3)[3],
  fill.mid = RColorBrewer::brewer.pal(name = 'PiYG', n = 3)[2],
  x.var = "alpha",
  y.var = "H2",
  fill.var = "mean.error",
  facet.x = "inference.method",
  facet.y = NA,
  is.oob.squish = F,
  is.real.scale = F,
  labeller = label_parsed) +
  labs(x = expression("Selection strength (" * alpha * ")"),
       y = expression("spVL heritability (" * italic('H'^2) * ")"),
       title = "A") +
  theme(
    text = element_text(size = 8),
    panel.border = element_rect(colour = "black", size = 0.5, fill = NA),
    strip.background = element_rect(colour = "black", size = 0.5))

show(rmse_plot)

ggsave(plot = rmse_plot,
       filename = "manuscript/figures/simulation_RMSE.png",
       width = 4.25,
       height = 2,  # maintain height as 2 for consistent font size in manuscript
       units = "in")
```

Simulation accuracy heatmap annotated with inferred POUMM values for HIV spVL
```{r}
rmse_plot_with_hiv_values <- rmse_plot +
  annotate(
    geom = "point",
    x = 8.046, y = 3.9)  # values are hand-calculated because the breaks scale is defined as seq(0.5, n.breaks - 0.5, 1) in plotting function

show(rmse_plot_with_hiv_values)

ggsave(plot = rmse_plot_with_hiv_values,
       filename = "manuscript/figures/simulation_RMSE_with_HIV_value_highlighted.png",
       width = 4.25,
       height = 2,  # maintain height as 2 for consistent font size in manuscript
       units = "in")
```

Simulation TPR heatmap
```{r}
data <- tidyr::gather(
  data = tpr_simulation_results,
  key = "inference.method",
  value = "p.value",
  h.plus.e.p.value, norm.z.p.value, h.OU.p.value, h.MWA.p.value)

# Calculate TPR for each simulation
sig.level <- correctSignificanceLevelForMultTesting(
  significance.level = 0.05,
  significance.correction = "Bonferroni",
  N.tests = 20)
data$TP <- ifelse(
  data$p.value < sig.level & data$is.causal.variant, T, F)
data$FP <- ifelse(
  data$p.value < sig.level & !(data$is.causal.variant), T, F)
data.TPR <- as.data.frame(
  data %>%
    dplyr::group_by(runID, inference.method, alpha, sigma, H2, var.z, p, delta) %>%
    dplyr::summarise(
      TPR = sum(TP, na.rm = T)/sum(is.causal.variant),
      FP = sum(FP), K = dplyr::n(),
      N.noncausal.variants = sum(!(is.causal.variant))))

# Summarize replicate simulations
data.summary <- as.data.frame(
  data.TPR %>%
    dplyr::group_by(inference.method, alpha, sigma, H2, var.z, p, delta) %>%
    dplyr::summarise(
      mean.TPR = mean(TPR),
      sd.TPR = sd(TPR),
      mean.FP = mean(FP),
      sd.FP = sd(FP),
      N.noncausal.variants = mean(N.noncausal.variants),
      N.sim = dplyr::n()))

# Create facet variable labels for plot
data.summary$inference.method <- factor(
  data.summary$inference.method,
  levels = c("h.plus.e.p.value", "h.MWA.p.value", "h.OU.p.value", "norm.z.p.value"),
  labels = c("e", "italic(hat(e))", "hat(e)[OU]", "'scaled spVL'"))

# Plot TPR 
tpr_plot <- makeTPRPlot(
  df = data.summary[
    data.summary$inference.method %in% 
      c("e", "italic(hat(e))", "'scaled spVL'"), ],
  is.MLE = T,
  midpoint = 0.5,
  fill.lo = RColorBrewer::brewer.pal(name = 'RdBu', n = 3)[3],
  fill.hi = RColorBrewer::brewer.pal(name = 'RdBu', n = 3)[1],
  fill.mid = RColorBrewer::brewer.pal(name = 'RdBu', n = 3)[2]) +
  facet_grid(. ~ inference.method,
             labeller = label_parsed) +
  labs(x = expression("Selection strength (" * alpha * ")"),
       y = expression("spVL heritability (" * 'H'^2 * ")"),
       title = "B") +
  theme(
    text = element_text(size = 8),
    panel.border = element_rect(colour = "black", size = 0.5, fill = NA),
    strip.background = element_rect(colour = "black", size = 0.5))

show(tpr_plot)

ggsave(plot = tpr_plot,
       filename = "manuscript/figures/simulation_TPR.png",
       width = 5.25,
       height = 2,  # maintain height as 2 for consistent font size in manuscript
       units = "in")
```
TPR heatmap annotated with inferred POUMM values for HIV spVL
```{r}
tpr_plot_with_hiv_values <- tpr_plot +
annotate(
geom = "point",
x = 8.046, y = 3.9)  # values are hand-calculated because the breaks scale is defined as seq(0.5, n.breaks - 0.5, 1) in plotting function

show(tpr_plot_with_hiv_values)

ggsave(plot = tpr_plot_with_hiv_values,
filename = "manuscript/figures/simulation_TPR_with_HIV_value_highlighted.png",
width = 5.25,
height = 2,  # maintain height as 2 for consistent font size in manuscript
units = "in")
```

Combined figure
```{r}
png(
  filename = "manuscript/figures/simulation_results.png",
  width = 5, height = 4, unit = "in", res = 300)
gridExtra::grid.arrange(
  rmse_plot_with_hiv_values, tpr_plot_with_hiv_values,
  nrow = 2,
  layout_matrix = rbind(c(NA, 1, 1, 1),
                        c(2, 2, 2, 2)))
dev.off()
```