### 2022-06-11 manuscript/hiv-gwas.tex
Update on Overleaf.

\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{lineno} % for line numbers
\usepackage{graphicx}
\usepackage{bm} % for bold math characters
\usepackage{booktabs} % for prettier hlines in table
\usepackage{subcaption} % for subfigures
\usepackage[affil-it]{authblk} % for footnote-style author affiliation
\usepackage{setspace} % for 1.5-spaced lines not including footnotes, floats
\usepackage{tikz} % for graphical model drawing
\usepackage{float} % to enforce figure positioning in subsection they're declared in
\usepackage{amsmath}
\usepackage[singlelinecheck=false]{caption} % to left-justify single-line table captions
\usepackage{array}
\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}} % for centered, width-specified columns 
\usepackage{makecell} % for multi-row table entries
\usepackage{tabularx} % for tables that don't spill past margin

\usepackage{url} % otherwise bibtex hates my references file
\usepackage[round,sort]{natbib}
\linespread{1.3}

% Tikz preamble from Michael Landis
\usepackage{amssymb} %maths
\usepackage{amsmath} %maths
%\usepackage{subfig} % can't be used together with subcaption package
\usepackage{skull}
\usetikzlibrary{positioning}
\usetikzlibrary{arrows}
\usetikzlibrary{fit}
\usetikzlibrary{calc}
\usetikzlibrary{automata}
\usetikzlibrary{decorations.markings}
\usetikzlibrary{decorations.pathreplacing} % for curly bracehttps://www.overleaf.com/project/6127a9ffb8327d4ab8b9fd4d
\usetikzlibrary{shapes.multipart}
\tikzset{>=latex}
\tikzstyle{snode}=[black,draw=black,line width=1.5pt,shape=circle,fill=white,minimum size=8mm]
\tikzstyle{obnode}=[black,draw=black,line width=1.5pt,shape=circle,fill=black!20!white,minimum size=8mm]
\tikzstyle{detnode}=[black,draw=black,line width=1.5pt,densely dotted,shape=circle,fill=white,minimum size=8mm]
\tikzstyle{constnode}=[black,draw=black,line width=1.5pt,shape=rectangle,fill=white,minimum size=8mm]
\tikzstyle{blnode}=[white,draw=black,line width=1pt,shape=circle,fill=black,minimum size=1mm,font=\scriptsize,inner sep=1pt]
\tikzstyle{ylnode}=[black,draw=black,line width=1pt,shape=circle,fill=yellow,minimum size=1mm,font=\scriptsize,inner sep=1pt]
\tikzstyle{taro}=[->,line width=2pt,color=black]
\tikzstyle{baro}=[<->,line width=2pt,color=black]
\tikzstyle{bline}=[line width=2pt,color=black]
\tikzstyle{dtaro}=[->,line width=2pt, densely dotted,color=black]
\tikzstyle{dline}=[line width=2pt, densely dotted,color=black]
\tikzstyle{smod}=[black, draw=black, line width=2pt, fill=white, shape=rectangle, rounded corners, minimum size=10mm, minimum width=20mm]
\tikzstyle{obmod}=[black, draw=black, line width=2pt, fill=black!20!white, shape=rectangle, rounded corners, minimum size=10mm, minimum width=20mm, minimum width=20mm]

% For proof
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}

% Set up supplemental figure numbering
\newcommand{\beginsupplement}{\setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}}

\begin{document}

% \begin{flushright} % for MBE submission
% Submission type: Article\\
% Section: Methods
% \end{flushright}

\begin{center}
	\Large{\textbf{A phylogeny-aware GWAS framework to correct for heritable pathogen effects on infectious disease traits}}
\end{center}
\vspace{4pt}

\begin{flushleft}
	Sarah Nadeau$^{1,2}$, Christian W. Thorball$^{3}$, Roger Kouyos$^{4,5}$, Huldrych F. Günthard$^{4,5}$, Jürg Böni$^{4}$, Sabine Yerly$^6$, Matthieu Perreau$^{7}$, Thomas Klimkait$^{8}$, Andri Rauch$^{9}$, Hans H. Hirsch$^{8,10,11}$, Matthias Cavassini$^{12}$, Pietro Vernazza$^{13}$, Enos Bernasconi$^{14}$, Jacques Fellay$^{2,3,15}$, Venelin Mitov$^{\dagger,1,2}$, Tanja Stadler$^{\dagger,\ast,1,2}$, and the Swiss HIV Cohort Study (SHCS)
\end{flushleft}

\noindent
$^{1}$Department of Biosystems Science and Engineering, ETH Zurich, Basel, Switzerland\\
$^{2}$Swiss Institute of Bioinformatics, Lausanne, Switzerland\\ 
$^{3}$Precision Medicine Unit, Lausanne University Hospital and University of Lausanne, Lausanne, Switzerland\\ 
$^{4}$Institute of Medical Virology, University of Zurich, Zurich, Switzerland\\ 
$^{5}$Division of Infectious Diseases and Hospital Epidemiology, University Hospital Zurich, University of Zurich, Zurich, Switzerland\\
$^{6}$Division of Infectious Diseases, Laboratory of Virology, Geneva University Hospital, Geneva, Switzerland\\
$^{7}$Division of Immunology and Allergy, University Hospital Lausanne, Lausanne, Switzerland\\
$^{8}$Department of Biomedicine, University of Basel, Basel, Switzerland\\
$^{9}$Department of Infectious Diseases, Bern University Hospital and University of Bern, Bern, Switzerland\\
$^{10}$Division of Clinical Virology, University Hospital Basel, Basel, Switzerland\\
$^{11}$Division of Infectious Diseases and Hospital Epidemiology, University Hospital Basel, Basel, Switzerland\\
$^{12}$Division of Infectious Diseases, University Hospital Lausanne, Lausanne, Switzerland\\
$^{13}$Division of Infectious Diseases, Cantonal Hospital St. Gallen, St. Gallen, Switzerland\\
$^{14}$Division of Infectious Diseases, Regional Hospital Lugano, Lugano, Switzerland\\
$^{15}$Global Health Institute, School of Life Sciences, École Polytechnique Fédérale de Lausanne, Lausanne, Switzerland\\
$\dagger$Co-last authors\\
$*$Corresponding author: tanja.stadler@bsse.ethz.ch

\newpage
\begin{linenumbers}
% \begin{doublespace}

\begin{abstract}
	Infectious diseases are particularly challenging for genome-wide association studies (GWAS) because genetic effects from two organisms (pathogen and host) can influence a trait. Traditional GWAS assume individual samples are independent observations. However, pathogen effects on an infectious disease trait can be heritable from donor to recipient in transmission chains. Thus, residuals in GWAS association tests for host genetic effects may not be independent due to shared pathogen ancestry. We propose a new method to estimate and remove heritable pathogen effects on a trait based on the pathogen phylogeny prior to host GWAS, thus restoring independence of samples. In simulations, we show this pre-processing step can increase GWAS power to detect truly associated host variants when pathogen effects are highly heritable, with strong phylogenetic correlations. When we apply our method to HIV-1 subtype B data from the Swiss HIV Cohort Study, we recover slightly weaker but qualitatively similar signals of association between spVL and human genetic variants in the \emph{CCR5} and major histocompatibility complex (MHC) gene regions compared to standard GWAS. Our simulation study confirms that based on the estimated heritability and selection parameters for HIV-1 subtype B spVL, standard GWAS are robust to pathogen effects. Our framework may improve GWAS for other diseases if pathogen effects are even more phylogenetically correlated amongst individuals in a cohort.
\end{abstract}

\section*{Introduction}

% What is GWAS and why is it important
A key goal of genome-wide association studies (GWAS) is to understand the genetic basis of phenotypic variation among individuals. In a typical GWAS, millions of genetic variants from across an organism's genome are screened for statistical association with a trait of interest. Ideally, this procedure identifies variants that are located in, or are in linkage disequilibrium with, alleles that directly affect the trait. If GWAS finds a variant strongly associated with a disease trait, the gene product may be a good drug target \citep{Okada2014}. Even if no single variant has a strong association, many small associations can be aggregated into a polygenic risk score to identify susceptible individuals \citep{Dudbridge2013}.

% Overview of relevant methods and why we don't use them
It is well-known that GWAS can be sensitive to confounding variables. Shared ancestry among individuals, especially between close relatives, can give rise to spurious genetic correlations with a trait. Corrections for these types of population structure are well-developed and widely accepted  \citep{Astle2009, Price2006}. More recently, analogous methods have been developed for microbial GWAS, where clonal reproduction exacerbates population structure \citep{Power2017}. Microbial GWAS-specific phylogenetic methods to account for population structure in microbial GWAS include explicitly testing for lineage-specific effects as in \citep{Earle2016} and modified association tests that account for phylogenetic relationships amongst samples as in \citep{collins_phylogenetic_nodate}. 
% \citep{Earle2016} simultaneously test for pathogen lineage and locus-level associations with a trait using a linear mixed model framework. \citep{collins_phylogenetic_nodate} test for locus-level associations that are significantly stronger than expected under a null model considering shared pathogen ancestry. 
These approaches are designed to quantify genetic effects from one organism on a trait.

In the infectious disease context, genetic effects from two organisms - the host and the pathogen - may affect an infectious disease trait. GWAS using paired host-pathogen genotype data have previously been done to elucidate the marginal and interaction effects of host and pathogen genetic variants. Methods to account for microbial population structure when testing for marginal host associations or host-pathogen interaction effects include adding the microbial kinship matrix as a random effect in a linear mixed model as in \citep{Wang2018Two-wayGenomes} and using principle components derived from either this matrix or the pathogen phylogeny as covariates in a linear model as in \citep{Naret2018}. These methods focus on capturing and accounting for correlations due to the pathogen phylogeny, without further investigating the nature of these correlations.
%However, the top principle components capture only some of the information from the full pathogen phylogeny. Furthermore, we would like to simultaneously address issues (i) and (ii).

In this work, we draw from the field of phylogenetic comparative methods to propose a new two-step framework to correct for pathogen population structure and satisfy the GWAS assumption of independent samples. It is a pre-processing step for GWAS that first fits an evolutionary model to trait data and the pathogen phylogeny to estimate the correlation structure of the trait due to pathogen effects. On the basis of this inference, it estimates and removes pathogen effects on the trait so that downstream GWAS association testing with host genetic variants can be performed as normal under the assumption of independent samples. The main advantage of this approach compared to the previously outlined methods to correct for pathogen population structure are that it generates additional information on the evolutionary dynamics of the trait in the pathogen population. It is also restricted to a pre-processing step, which allows use of existing highly optimized tools to perform GWAS association tests under a variety of models. This framework relies on paired pathogen-host genotyping and is envisioned specifically for continuous-valued traits that are highly heritable from infection partner to infection partner.
% We hypothesis this should increase GWAS power for several reasons. For one, pathogen effects may be confounding or extraneous variables that produce distorted or spurious associations between host genetic variants and a disease trait in host GWAS. 
We hypothesized that our approach should improve GWAS power to identify host genetic variants broadly associated with disease traits.
%This may occur if the virus affects the disease trait (the dependent variable in host GWAS) or if it affects both the disease trait and the host genotype (the independent variable in host GWAS).
%For example, stratification in both the host and pathogen populations may yield spurious correlations in host GWAS that are really due to independent pathogen factors. 
% Then, co-evolution between a pathogen and its host may yield distorted estimates for host genetic effects that are not broadly protective against all pathogen strains. Finally, trait values from a population are not truly independent samples, given that hosts closer in the transmission chain have more similar pathogen strains and therefore more similar trait values.

% % Why is spVL and important trait
% For HIV, GWAS have used a trait called set point viral load (spVL) to identify human variants associated with the severity of disease course. spVL is generally defined to be the average concentration of viral RNA copies in host plasma during the asymptomatic phase of infection in the absence of treatment (see e.g. \cite{Alizon2010}). In untreated individuals, spVL is predictive of the duration of asymptomatic infection \citep{Mellors1996} and infectiousness \citep{Quinn2000}. If viral load can be reduced to undetectable levels, an individual is effectively uninfectious and the risk of disease progression is massively reduced \citep{HHSARTGuidelines2019}. Notably, spVL varies by orders of magnitude between individuals \citep{Mellors1996}. Thus, spVL measurements point to a wide range in natural HIV control amongst individuals.

% % How do we measure the extent to which genetic factors determine spVL
% To understand HIV pathogenicity, it is important to understand to what extent spVL is determined by host genetic factors as opposed to pathogen genetic factors. Heritability is a key measure of how genetically-determined a trait is. On the host side, narrow-sense heritability, the fraction of total trait variance explained by additive genetic effects, is a typical output of GWAS. 
% % How heritable is spVL (human side)
% Several GWAS have measured the narrow-sense heritability of spVL and identified associated host genetic variants \citep{Bartha2013, Dalmasso2008, Fellay2007, Pereyra2010, Fellay2009, Pelak2010, VanManen2011, McLaren2012}. The largest study to-date by \citet{McLaren2015} estimated that additive human genetic effects account for approximately 25\% of the variation in spVL amongst infected individuals. All but 5\% of this was attributed to two regions in the human genome, the major histocompatibility complex (MHC) and C-C motif chemokine receptor 5 (\emph{CCR5}). Both associations are biologically relevant: the MHC encodes proteins that present viral epitopes at the cell surface and \emph{CCR5} encodes a co-receptor for HIV-1 cell entry. In other words, MHC proteins match bits of the virus like puzzle pieces and display these to signal that a cell is infected. CCR5 proteins help the virus infect target cells.

% How heritable is spVL (viral side)
% On the pathogen side, broad-sense heritability $H^2$ is typically measured. This is the fraction of total trait variance due to any factors ``inherited'' from infection partner to infection partner (the most obvious being pathogen genetic factors). Broad-sense heritability is commonly measured by quantifying how correlated a trait is between infection partners or how phylogenetically structured a trait is based on a reconstructed pathogen phylogeny. For HIV spVL, estimates differ depending on the methods employed and the cohort studied (see \citet{Mitov2018} for a discussion of this uncertainty). Using phylogenetic methods on large UK and Swiss cohorts, \citet{Mitov2018} and \citet{Bertels2018} estimated that pathogen factors account for 21\% - 29\% of the variation in spVL amongst infected individuals. However, variation in the MHC is known to exert strong selective pressure on the virus \citep{Kloverpris2016, Nguyen2021SystematicLink}. If the virus can change its ``puzzle piece'' shape to escape MHC-presentation, infected cells can go undetected. This means that MHC variants affect spVL largely via selection on the virus \citep{Bartha2017}. In summary, human genetic factors play a role in determining spVL, but these effects may be due to interaction with specific viral genetic variants.

% % The gap
% Most of the GWAS for human genetic determinants of spVL conducted so far \citep{Dalmasso2008, Fellay2007, Pereyra2010, Fellay2009, Pelak2010, VanManen2011, McLaren2012, McLaren2015} do not explicitly consider any viral effect on spVL. In these GWAS, viral genetic effects are lumped in with residual variance due to other, non-genetic factors. This has several potential negative consequences. 
% n GWAS, hosts are typically selected to not share recent ancestry. Further, shared environmental variables are included as covariates to avoid spurious associations. We argue similar considerations must be taken
% for pathogen effects that heritable from one transmission partner to another \citep{Leventhal2016}.

% Introduction to our method
In the following, we describe the evolutionary model for heritable, continuous-valued infectious disease traits upon which our method is based. We then derive a maximum likelihood estimate for the pathogen part of a trait under this model, and describe a new infectious disease GWAS framework using this information. In simulations, we show that this framework can improve GWAS power to detect host genetic variants that affect disease traits. Finally, we apply our framework to paired host-pathogen genotyping data from the Swiss HIV Cohort Study (SHCS) and a previously studied \emph{Arabidosis thaliana}-\emph{Xanthomonas arboricola} pathosystem. We show that associations with set-point viral load (spVL) and quantitative disease resistance (QDR) traits, respectively, are robust to a correction for pathogen effects.

\section*{New Approaches}
\subsection*{A statistical model for heritable, continuous-valued infectious disease traits}

Variation in infectious disease traits like viral load or infection severity can come from several sources. These include host genetic factors, pathogen genetic factors, interaction effects between the host and the pathogen, or non-genetic factors like healthcare quality or temperature. GWAS typically stratify samples or include covariates to correct for host genetic factors or non-genetic factors that may be correlated with a trait value. This leaves pathogen genetic factors as a remaining source of correlation, since close transmission partners may be infected with very similar pathogen strains. Broad-sense heritability $H^2$ quantifies the fraction of total trait variance that is ``inherited'' from infection partner to infection partner. To characterize $H^2$ and the heritable and non-heritable factors that determine infectious disease traits, we use a phylogenetic mixed model (PMM) \citep{Housworth2004}. PMMs assume continuous traits are the sum of independent heritable and non-heritable parts. In the infectious disease GWAS case, we assume the heritable part comprises pathogen genetic factors and all other factors are non-heritable. The heritable pathogen part is modeled by a random process occurring in continuous time along the branches of the pathogen phylogeny, as in Figure \ref{fig:spVLModel}A. The non-heritable part is modeled as Gaussian noise added to sampled individuals at the tips of the phylogeny.

\begin{figure}[H]
	\begin{center}
	\includegraphics[width = \linewidth]{figures/simulation_setup.png}
	\caption{A high-level schematic of our POUMM-based simulation framework in the context of HIV-1 spVL. (A) shows how the viral effects on spVL evolve along the viral phylogeny according to an Ornstein-Uhlenbeck process. (B) shows how human host genetic effects are the sum of independent effects from several causal variants. Each variant can be present in 0, 1, or 2 copies. Half the variants have a positive effect of size $\delta$ and half have a negative effect of size $\delta$. (C) shows how other environmental effects are independently drawn from a Gaussian distribution centered at 0. These three effects sum to the trait value for each simulated individual.}
		\label{fig:spVLModel}
	\end{center}
\end{figure}

PMMs have previously been applied to the study of infectious disease traits using two different types of random processes to model trait evolution. The Brownian Motion (BM) process assumes unbounded trait values, i.e. the trait can attain any value. The Ornstein-Uhlenbeck (OU) process assumes trait values fluctuate around an optimal value, i.e. extreme trait values are unlikely. Here, we assume the more flexible OU process as it encompasses a wider variety of evolutionary scenarios. For example, \citet{Mitov2018} and \citet{Bertels2018} previously showed the OU process has higher statistical support for HIV-1 spVL. This makes sense given that spVL is likely under stabilizing selection to maximize viral transmission potential \citep{Fraser2014}. The full model is called the phylogenetic Ornstein-Uhlenbeck mixed model (POUMM) and is described in detail by \citet{Mitov2018}. Here, we review the main points relevant to our method.

Under the POUMM, the trait $z$ is the sum of heritable genetic effects $g$, i.e. due to the pathogen, and non-heritable ``environmental'' effects $\epsilon$, i.e. host genetic effects and other environmental or interaction effects:

\begin{equation}
    z = g + \epsilon
\end{equation}
$g$ is a pathogen trait that evolves along the phylogeny according to an OU process. The OU process is defined by a stochastic differential equation with two terms. The first term represents a deterministic pull towards an optimal trait value and the second term represents stochastic fluctuations modelled by Brownian motion \citep{Butler2004}:

\begin{align}
\begin{split}
	&dg(t) = \alpha[\theta - g(t)]dt + \sigma dW_t \\
	&g(0) = g_0 
	\label{eq:OUprocess}
\end{split}
\end{align}

Here the parameter $\alpha$ represents selection strength towards an evolutionarily optimal value represented by parameter $\theta$. The parameter $\sigma$ measures the intensity of stochastic fluctuations in the evolutionary process. Finally, $dW_t$ is the Wiener process underlying Brownian motion. The OU process is a Gaussian process, meaning that $g(t)$ is a Gaussian random variable. Assuming $g(t)$ starts at initial value $g_0$ at time $t = 0$ at the root of the phylogeny, we can write the expectation for $g(t)$ at time $t$:

\begin{equation}
   E[g(t)] = g_0e^{-\alpha t} + (1 - e^{-\alpha t})\theta \label{eq:OUmean}
\end{equation}
and the variance in $g(t)$ if we were to repeat the random evolutionary process many times \citep{Butler2004}:
\begin{equation}
    Var[g(t)] = \frac{\sigma^2}{2\alpha}(1 - e^{-2\alpha t}) \label{eq:OUVar}
\end{equation}

$g$ evolves independently in descendent lineages after a divergence event in the phylogeny. The covariance between $g(t)$ in a lineage $i$ at time $t_i$ and another lineage $j$ at time $t_j$, $Cov\big(g_{i}(t_i), g_{j}(t_j)\big)$, increases with the amount of time between $t_0$ and the divergence of the two lineages, $t_{0(ij)}$, and decreases with the total amount of time the lineages evolve independently, $d_{ij}$ \citep{Butler2004}:

\begin{equation}
	Cov\big(g_{i}(t_i), g_{j}(t_j)\big) = \frac{\sigma^2}{2\alpha}[e^{-\alpha d_{ij}}(1 - e^{-2\alpha t_{0(ij)}})]
	\label{eq:OUcov}
\end{equation}
Next, we remember that $\epsilon$ is the non-heritable part of the trait. $\epsilon$ is modeled as a Gaussian random variable that is time- and phylogeny-independent. The expectation of $\epsilon$ is 0, meaning environmental effects are equally likely to raise or lower the trait from the pathogen-determined level. The parameter $\sigma_\epsilon^2$ measures the between-host variance of the environmental effect.

\begin{align}
\begin{split}
	E(\epsilon) = 0 \\
	Var(\epsilon) = \sigma^2_\epsilon
\end{split}
\end{align}

Finally, broad-sense trait heritability can be calculated as the fraction of total trait variance that is heritable:

\begin{equation}
	H^2_{\bar{t}} = \frac{Var[g(t)]}{Var[g(t)] + Var(\epsilon)} = \frac{\frac{\sigma^2}{2\alpha}(1 - e^{-2\alpha t})}{\frac{\sigma^2}{2\alpha}(1 - e^{-2\alpha t}) + \sigma^2_\epsilon}
	\label{eq:POUMM-H2}
\end{equation}

\subsection*{Teasing apart pathogen and non-pathogen effects on a trait}

Given the assumptions of the POUMM, we can estimate a heritable pathogen effect on a trait and a corresponding non-heritable, host and environmental effect. Here, we derive a maximum-likelihood estimate for these values for individuals in a GWAS cohort, given measured trait values and a pathogen phylogeny linking the infecting strains.

Let $\bm{g}(\bm{t})$ be a vector of $g$ values, one for each individual in the cohort. $\bm{t}$ are the sampling times of each individual relative to the root of the phylogeny. To simplify notation, we omit the $\bm{t}$ from here on. $\bm{g}$ is a realization of a Gaussian random vector $\bm{G} \sim \mathcal{N}\big(\bm{\mu}_{OU}, \boldsymbol{\Sigma}_{OU}\big)$. The expectation $\bm{\mu}_{OU}$ is defined by equation \ref{eq:OUmean}, the diagonal elements of the covariance matrix $\boldsymbol{\Sigma}_{OU}$ are defined by equation \ref{eq:OUVar}, and the off-diagonal elements of $\boldsymbol{\Sigma}_{OU}$ by equation \ref{eq:OUcov}. Similarly, let $\bm{\epsilon}$ be a vector of the environmental part of the trait for each individual.  $\bm{\epsilon}$ is a realization of a Gaussian random vector $\bm{\mathcal{E}} \sim \mathcal{N}\big(\bm{0}, \boldsymbol{\Sigma}_\mathcal{E}\big)$, where $\boldsymbol{\Sigma}_\mathcal{E}$ is a diagonal matrix with diagonal elements equal to $\sigma^2_\epsilon$.

Considering that $\bm{G}$ and $\bm{\mathcal{E}}$ are independent random vectors and that their realizations $\bm{g}$ and $\bm{\epsilon}$ must sum together to equal the observed trait values $\bm{z}$, we can write the following proportionality for the joint probability density of $\bm{g}$ and $\bm{\epsilon}$:

\begin{equation}
	f\big(\bm{g}, \bm{\epsilon}\big) \propto \mathcal{N}\big(\bm{g}; \bm{\mu}_{G}, \boldsymbol{\Sigma}_G\big)
	\label{eq:pdfGprop}
\end{equation}

where the expected value of $\bm{g}$ and the covariance matrix $\boldsymbol{\Sigma}_G$ are defined as:

\begin{align}
	Exp(\bm{g}) = \bm{\mu}_{G} &=  \boldsymbol{\Sigma}_G\big(\boldsymbol{\Sigma}_{OU}^{-1}\bm{\mu}_{OU} + \boldsymbol{\Sigma}_\mathcal{E}^{-1} \bm{z}\big) \label{eq:MuG}\\
	\boldsymbol{\Sigma}_G &= \big(\boldsymbol{\Sigma}_{OU}^{-1} + \boldsymbol{\Sigma}_\mathcal{E}^{-1}\big)^{-1} \label{eq:SigmaG}
\end{align}

\begin{proof}
	\begin{align}\label{eq:pdfG}
	\begin{split}
		f\big(\bm{g},\ \bm{\epsilon}\big) &= f\big(\bm{g}|\ \bm{\epsilon}\big) \times f\big(\bm{\epsilon}\big) \\
	&= f\big(\bm{g}\big) \times f\big(\bm{\epsilon}\big) \\
	&= \mathcal{N}\big(\bm{g};\ \bm{\mu}_{OU}, \mathbf{\Sigma}_{OU}\big) \times \mathcal{N}\big(\bm{\epsilon};\ \bm{0}, \mathbf{\Sigma}_\mathcal{E}\big) \\
	&= \mathcal{N}\big(\bm{g};\ \bm{\mu}_{OU}, \mathbf{\Sigma}_{OU}\big) \times \mathcal{N}\big(\bm{z} - \bm{g};\ \bm{0}, \mathbf{\Sigma}_\mathcal{E}\big) \\
	&= \mathcal{N}\big(\bm{g};\ \bm{\mu}_{OU}, \mathbf{\Sigma}_{OU}\big) \times \mathcal{N}\big(\bm{g};\ \bm{z}, \mathbf{\Sigma}_\mathcal{E}\big)
	\end{split}
\end{align}

	Equations \ref{eq:MuG} and \ref{eq:SigmaG} follow from eq. \ref{eq:pdfG} and eq. 371, p. 42, section 8.1.8 ``Product of Gaussian densities'' in \citet{Petersen2012}.
\end{proof}

Importantly, equation \ref{eq:MuG} is the maximum likelihood estimate for $\bm{g}_v$, the pathogen effect on the trait, taking into account all available information - measured trait values, the pathogen phylogeny, and inferred POUMM parameters. This estimator is an inverse-variance weighted average of measured trait ($\bm{z}$) and information from the POUMM evolutionary model ($\bm{\mu}_{OU}$). In other words, $g$ will be closer to the measured trait value if the trait is not very heritable. If the trait is highly heritable, $g$ will be closer to the expected value under the POUMM, i.e. take more information from the phylogenetic relationships between infecting strains.

Given the estimator we just derived for $\bm{g}$, we can now estimate $\bm{\epsilon}$, the trait value $without$ pathogen effects:

\begin{equation}
	\hat{\bm{\epsilon}} = \bm{z} - Exp(\bm{g})
	\label{eq:EHat}
\end{equation}

We will use this value to try to improve upon standard GWAS methods in infectious disease.

\subsection*{A POUMM-based GWAS framework for infectious disease}

We propose to improve standard GWAS for infectious diseases by estimating and removing trait variability due to pathogen effects. Our new framework is as follows:

\begin{enumerate}
	\item Sample paired host genotypes, pathogen genome sequences, and trait values from a cohort.
	\item Construct a pathogen phylogeny using the pathogen genome sequences.
	\item Estimate the parameters of the POUMM based on the trait values and the pathogen phylogeny. This can be done with e.g. the R package POUMM \citep{Mitov2017a-POUMM}.
	\item Generate maximum-likelihood estimates for the pathogen and corresponding non-pathogen effects on the trait using equations \ref{eq:MuG} and \ref{eq:EHat}.
	\item Perform GWAS with only the non-pathogen effects on the trait as the response variable.
\end{enumerate}

\section*{Results}

\subsection*{Simulation study}

To test the theoretical best-case performance of our method, we simulated data under the POUMM and applied our framework to the simulated data. We parameterized our simulation scheme with the time-scale and other parameters of an HIV-1 outbreak in mind, with spVL as the trait of interest.

We first simulated a phylogeny of 500 tips with exponentially distributed branch lengths and mean root-to-tip time of 0.14 substitutions per site per year as in \cite{Hodcroft2014}. Then, we simulated pathogen trait values $g$ along this phylogeny using the POUMM package in R \citep{Mitov2017a-POUMM}. For the simulation, we considered a range of pathogen heritability parameter values $H^2$, from 15 to 75\%, and a range of selection strength parameters values $\alpha$, from 0.1 to 60 time$^{-1}$. The intensity of stochastic fluctuations parameter $\sigma$ was determined based $H^2$ and $\alpha$ (a re-arrangement of equation \ref{eq:POUMM-H2} given the partition of trait variance described below). These values ranged from XX to XX, depending on alpha. We recorded the simulated pathogen part of the trait value for each tip in the phylogeny.

We paired each tip's simulated pathogen trait value with a simulated host trait value. Simulated hosts had 100 genome positions. We sampled alleles (0, 1, or 2) for each position from a binomial distribution with probability 0.13. 10 random positions had an effect size of 0.2 on the trait and 10 had an effect size of -0.2. The other 80 positions had no effect on the trait. This parameterization produced roughly normally distributed host trait values centered at 0 with variance equal to 25\% of the total trait variance, which we constrained to 0.73 based on the variance in log spVL values measured by \cite{Mitov2018}. We used 25\% host heritability for spVL based on \cite{McLaren2015}.

Finally, we sampled an additional random environmental effect for each tip from a normal distribution centered at 0. The variance of this distribution was scaled based on the pathogen heritability of the trait, from 0 (no affect) in the scenario with 75\% pathogen heritability and 25\% host heritability to 0.44 in the scenario with 15\% pathogen heritability and 25\% host heritability. Figure \ref{fig:sim-design} shows a high-level schematic of this simulation framework and Table \ref{tab:sim-params} gives the value or expression for each parameter.

\subsubsection*{Estimator accuracy}

First, we evaluated how well our method estimated the additive host genetic effects from the simulated data. Additive host genetic effects represent an ideal (albeit unattainable) baseline for infectious disease GWAS. Figure \ref{fig:simulationResults}A shows that our method incorporating phylogenetic information can more accurately estimate these value compared to the trait value. To ensure a fair comparison, we scaled trait values to have the same mean, zero, as host genetic effects so as not to bias the root mean squared error (RMSE) by a constant factor. As shown in the supplemental material, we can calculate the expected RMSE using the scaled trait value across scenarios in our simulation scheme because the variance in the trait due to viral genetic effects and environmental effects is fixed. Thus, we expect the RMSE using the scaled trait value to be 0.74 across all simulation scenarios. By incorporating phylogenetic information, we can improve upon this error in scenarios where the trait is highly heritable, under low selection pressure, and with relatively moderate stochastic fluctuations compared to outbreak duration. This is because high heritability means the trait value is highly pathogen-dependent. Then, when the trait is under weak selection, the pathogen effects can drift far from the long-term optimum and stochastic fluctuations are low to maintain the same heritability. Thus, our method performs well when an infectious disease trait is highly heritable and trait values are strongly correlated amongst close transmission partners.

\subsubsection*{Theoretical GWAS improvement}

Next, we characterized the evolutionary scenarios under which our framework can actually improve GWAS power. We used the true positive rate (TPR) to evaluate the fraction of simulated causal host genetic variants we could recover as being significantly associated with the trait. We performed three different GWAS for each simulated dataset: the first represents an ideal in which we can exactly know and remove pathogen effects from trait values, the second is using our method to estimate this value and remove it, and the third represents a standard GWAS using the scaled trait value. Figure \ref{fig:simulationResults}B shows that our framework can improve the TPR in simulated scenarios where selection strength $<$ 10 time$^{-1}$ and heritability $>$ 45\%. If we were able to perfectly estimate and remove pathogen effects from a trait, the TPR would increase across all values of selection strength so long as the trait is more than marginally heritable. We estimate approximately 25\% to be the heritability threshold above which GWAS power is negatively impacted by pathogen effects. In summary, we show it is theoretically possible to improve GWAS power for heritable infectious disease traits by estimating and removing pathogen effects using information from the pathogen phylogeny.

% More accurate estimation of the non-viral part of spVL does not guarantee improved GWAS power. For example, if the viral part of spVL were constant across all viruses, association results (significance and effect size) would not be effected by a viral contribution to the trait. 
% Under our simulation set-up a standard GWAS would have $\approx$ 30\% power to detect a host variant explaining 1.25\% of the variability in spVL. This matches simulation results (figure \ref{fig:simulationResults}B, scaled spVL).

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=\linewidth]{figures/simulation_results}
		\caption{Results from the simulation study. We simulated host, pathogen, and environmental effects on a trait under the POUMM with different heritability (y-axis) and selection strength (x-axis) parameters. For each simulated dataset, we applied our method to estimate the non-pathogen effects and performed GWAS with these values. (A) shows that our method (left) can generate more accurate estimates of additive host genetic effects than the trait value, scaled by its mean (right). (B) shows how GWAS power can improve given the true, simulated non-pathogen effect on spVL (left) and using our estimate for this value (middle) compared to using the scaled trait value (right). Each tile's color corresponds to the average value across 20 simulated datasets of 500 samples. The black point represents our estimates for the heritability and selection strength of spVL based on Swiss HIV Cohort Study data. The empty points show simulated scenarios that are presented in more detail in Figure \ref{fig:simulationTraitExample}. RMSE = Root mean square error, TPR = True positive rate.}
		\label{fig:simulationResults}
	\end{center}
\end{figure}

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=0.7\linewidth]{figures/example_simulation_trait_values}
		\caption{Example simulated data from two evolutionary scenarios where (left) a phylogenetic correction to trait values does not improve GWAS power and (right) where it does. These examples correspond to the empty points in Figure \ref{fig:simulationResults}. Apparent is that the phylogenetically-corrected trait values (y-axis) correspond better with the true host part of the trait when trait heritability due to the pathogen is higher (right side). Selection strength was fixed to 0.1 time $^{-1}$ for both scenarios and all other parameters were fixed as in the full simulation study. }
		\label{fig:simulationTraitExample}
	\end{center}
\end{figure}

\subsection*{GWAS on the Swiss HIV Cohort}

Finally, we applied our framework to empirical data from the Swiss HIV Cohort Study (SHCS). We used data collected from 1,392 individuals in Switzerland infected with HIV-1 subtype B between 1994 and 2018. The SHCS provided viral load measurements, $pol$ gene sequences, and human genotype data for these individuals. We followed the framework outlined above to estimate the pathogen and non-pathogen effects on spVL for the cohort from these data. Figure \ref{fig:spvl-calc-comparison} shows the calculated (total) spVL values, which vary between approximately 1 and 6 log copies/mL in the cohort.  Figure \ref{fig:spvl-on-tree} shows that this trait is not strongly phylogenetically structured in the cohort, despite high heritability. Finally, figure \ref{fig:trait-comparison} shows that the estimated non-pathogen effects on spVL correlate quite strongly with total spVL. We estimated spVL heritability in this cohort to be 45\% (95\% highest posterior density, HPD, 24 - 67\%) and selection strength to be 58 time$^{-1}$ (95\% HPD 19 - 95) (Figure \ref{fig:poumm-parameters}, Table \ref{tab:POUMM-parameters-our-estimates}). To put these values into the context of our simulation study, they are shown as black points on Figure \ref{fig:simulationResults}.

We compared our proposed GWAS framework with a more standard approach by performing two different GWAS on the same SHCS human genotypes. In the ``GWAS with standard trait value'' we used the total trait value, our calculated spVL values, as the GWAS response variable. In the ``GWAS with estimated non-pathogen part of trait'' we used our estimates for the non-pathogen effects on spVL. Figure \ref{fig:gwas-results}A shows that results are qualitatively similar between the two GWAS. Q-Q plots show the distribution of p-values are very similar as well (Figure \ref{fig:qq-plots}). Figure \ref{fig:gwas-results}B shows how the strength of association changed for some variants in the MHC and \emph{CCR5} regions. Taking into account phylogenetic information slightly decreased association strength for most variants in the \emph{CCR5} region. Association strength increased for some variants in the MHC, for example, SNP rs9265880 had the greatest increase in significance in the MHC region, from a p-value of $3.5 \times 10^{-07}$ to $7.7 \times 10^{-09}$. However, the top-associated variants in the MHC and \emph{CCR5} regions were consistent regardless of the GWAS response variable used (Table \ref{tab:gwas-top-associations}). Finally Table \ref{tab:comp-gwas-mclaren-snps} shows how our GWAS results compare for the two top-associated SNPs identified by \citet{McLaren2015}. In summary, there are no clear patterns that point to new regions of association in the human genome with spVL when we take into account the pathogen phylogeny.

\begin{figure}[H]
	\begin{center}
		\includegraphics[width=\linewidth]{figures/gwas_results}
		\caption{Results from comparative GWAS. (A) shows association p-values for the same host variants from the SHCS cohort in GWAS with two different response variables. On the left, we used unmodified (total) spVL values. On the right, we used our estimates for the non-pathogen effects on spVL. The alternating shades correspond to different chromosomes. (B) compares the strength of association for variants in the \emph{CCR5} and MHC regions between the two GWAS (positions 45.4 - 47Mb on chromosome 3 and 29.5 - 33.5Mb on chromosome 6 for the \emph{CCR5} and MHC, respectively). Base positions are with reference to genome build GRCh37. The color of each point represents the difference in -log$_{10}$ p-value between the two GWAS. Red means taking into account phylogenetic information decreased the strength of association and blue means it increased it. The dashed lines show genome-wide significance at p = $5 \times 10^{-8}$.}
		\label{fig:gwas-results}
	\end{center}
\end{figure}

\begin{table}[H]
\centering
\caption{Top association results from \citet{McLaren2015} compared to results from this study. Results from this study are for host variants from the SHCS in GWAS with two different response variables. ``Standard trait value'' means we used the unmodified (total) spVL value and ``Estimated non-pathogen part of trait'' means we used our estimates for the non-pathogen effects on spVL.}
\begin{tabularx}{\linewidth}{p{1cm}p{1.8cm}p{2.4cm}p{1.8cm}p{2cm}p{1.8cm}p{2cm}} % {lrrrrrr} {llclcccc}
\hline
& & \makecell{McLaren \\ et al.}  & \multicolumn{2}{l}{\makecell{Standard \\ trait value}} & \multicolumn{2}{l}{\makecell{Estimated non-pathogen \\ part of trait}} \\ 
  \hline
 Region & Variant & p-value & Effect size & p-value & Effect size & p-value \\
  \hline
    MHC & rs59440261 & $2.0 \times 10^{-83}$ & -0.4 & $3.3 \times 10^{-11}$ & -0.22 & $2.6 \times 10^{-10}$ \\ 
    \emph{CCR5} & rs1015164 & $1.5 \times 10^{-19}$ & 0.15 & $7.5 \times 10^{-7}$ & 0.078 & $8.5 \times 10^{-6}$ \\ 
   \hline
\end{tabularx}
\label{tab:comp-gwas-mclaren-snps}
\end{table}

\subsection*{GWAS on the \emph{A. thaliana}-\emph{X. arboricola} pathosystem}

\section*{Discussion}
% State the importance of your findings
In this paper, we presented a new phylogeny-aware GWAS framework to correct for heritable pathogen effects on infectious disease traits. By using information from the pathogen phylogeny, we show it is possible to improve GWAS power to detect host genetic variants associated with a disease trait. This should help us better understand which host factors are protective against a disease versus which increase susceptibility or disease severity.

% Contextualize the study & address limitations
Our method relies on the POUMM, a model of continuous trait evolution that accounts for heritable and non-heritable effects on a trait, as well as selection. Using this model, we estimated HIV-1 spVL heritability to be 45\% (95\% HPD 24 - 67\%) in the Swiss HIV Cohort Study. Compared to previous studies, this estimate is at the higher end (see \citet{Mitov2018} and references therein). Also using the POUMM, \citet{Bertels2018} estimated a spVL heritability of 29\% (N = 2014, CI 12 - 46\%) from the same cohort and \citet{Blanquart2017} estimated 31\% (N = 2028, CI 15 - 43\%) from a pan-European cohort. We note that our sample size (N = 1493 individuals) is smaller than in these other studies. This might be because we restricted samples based on having $pol$ gene sequences with at least 750 non-ambiguous bases. Our aim was to reconstruct a high-quality phylogeny, since the POUMM does not account for phylogenetic uncertainty and the POUMM parameter estimates are key to our downstream trait-correction method. Although our heritability estimate is rather high, the confidence interval largely overlaps that of other studies and we note that estimating heritability per se was not our primary focus.

Instead, the main novelty of our approach was to correct the spVL trait prior to performing the GWAS, thereby estimating and removing pathogen effects. In simulations, we show that when trait heritability amongst infection partners is greater than approximately 25\%, GWAS power to detect host genetic variants associated with the same trait is reduced. Our method can correct for this effect in certain evolutionary scenarios by using information from the full pathogen phylogeny.

Based on our simulation results, our method is anticipated to be very useful for disease traits that are highly heritable from donor to recipient and  maintain a high correlation between sampled individuals. In simulations, we showed this is the case when heritability is high, selection strength is low, and trait values are not subject to strong stochastic fluctuations. So, cohort-level, phylogenetically structured differences in the measured trait value are necessary for our approach to outperform state of the art methods. Co-infection, or any other non-genetic factor that increases variability in the trait of interest, would reduce the effectiveness of this method. The more environmental ``noise'' in the trait value, the less signal there is to tease apart pathogen effects. There is also less signal for host genetic effects, so GWAS in general suffer when there are confounding factors affecting a trait of interest. In this case, we would recommend adding co-infection status as a covariate in the GWAS association test.

Given our estimates for the heritability of spVL and the selection strength on this trait using Swiss HIV cohort data, our simulation results reveal that we cannot expect a significant improvement in GWAS power for human genetic determinants of spVL (Figure \ref{fig:simulationResults}). Our method slightly decreases p-values for variants in $CCR5$ and slightly decreases some and increases other p-values for variants in the MHC (Figure \ref{fig:gwas-results}B). Simulations show we shouldn't expect a net p-value decrease, but our simulations represent an ideal scenario since we simulate under the POUMM. In real life, un-modeled evolutionary pressures like drug treatment and host-specific HLA alleles might cause the reduced p-values. However, the overall picture is consistent between the two GWAS (Figure \ref{fig:gwas-results}A). Therefore, we conclude that GWAS for host determinants of HIV-1 subtype B spVL is robust to our correction for pathogen effects.

Our method is convenient for GWAS because it is simply a pre-processing step that produces an alternate response variable for GWAS association tests. It is still possible to use previously developed, well-documented, and fast tools for the actual association testing (we used PLINK \citep{Chang2015}). The method relies on the freely available R package POUMM \citep{Mitov2017a-POUMM}, which scales to trees of up to 10,000 tips \citep{Mitov2019}. All code for the simulations and HIV spVL analysis presented in this study is available on the project GitHub at https://github.com/cevo-public/POUMM-GWAS. Future applications of our method might investigate other clinically significant disease traits and outcomes that are affected by both host and pathogen genetic factors, for instance Hepatitis B Virus-related hepatocellular carcinoma \citep{An2018HostCarcinoma}, Hepatitis C treatment success \citep{Ansari2017Genome-to-genomeVirus}, and susceptibility to or severity of certain bacterial infections, e.g. \cite{Messina2016ImpactInfections, Donnenberg2015BacterialStudies}.

% Summarize the paper
In summary, we argue that infectious disease GWAS should take the pathogen phylogeny into account when searching for host determinants of a disease trait. We give a practical threshold for identifying when GWAS suffer from pathogen effects (heritability of the trait amongst infection partners $>$ 25\%) and provide a method that can help in scenarios where trait values are highly heritable and phylogenetically-structured amongst members of a cohort.

\section*{Materials and Methods}

\subsection*{Simulation model}

Whenever possible, we tried to parameterize our simulation model for spVL using empirical data. We set the total variance in spVL to 0.73 log copies$^2$ mL$^{-2}$ based on UK cohort data \citep{Mitov2018}. Other studies have estimated slightly lower values though (Table \ref{tab:spVLmeasurements}). After allotting 25\% of this variance to the host part of spVL $h$ based on results by \citet{McLaren2015}, we partitioned the remaining variance between the viral part $g$ and the environmental part $\epsilon$ in different ratios to assess estimator performance across a range of spVL heritabilities. $h$ was simulated as the sum of contributions from 20 causal host genetic variants, 10 of which had an effect size of 0.2 log copies mL$^{-1}$ and 10 of which had an effect size of -0.2 log copies mL$^{-1}$. Host genetic variants were generated from a binomial distribution with probability $p$ calculated such that $h$ had the appropriate variance (see Table \ref{tab:sim-params}). We generated a random viral phylogeny with branch lengths on the same time scale as a previously inferred UK cohort HIV tree \citep{Hodcroft2014} using the R package ape \citep{Paradis2018}. $g$ was simulated by running an OU process along the phylogeny using the R package POUMM \citep{Mitov2017a-POUMM} and sampling values at the tips. For the OU parameters $\theta$ and $g_{0}$ we used 4.5 log copies mL$^{-1}$ based on fitting the same model to SCHS data (Table \ref{tab:POUMM-parameters-our-estimates}). This is similar to values previously inferred for HIV (Table \ref{tab:POUMMparams-prev-estimated}). To assess our estimator's performance under a range of evolutionary scenarios, we co-varied the OU parameters for selection strength, $\alpha$, and intensity of random fluctuations, $\sigma$, so that different proportions of the variability in $g$ were attributable to selection and drift, respectively. Finally, the environmental component of spVL $\epsilon$ was generated from a normal distribution with mean 0.  For a full graphical model representation of the simulation scheme, see Figure \ref{fig:sim-design}.

\begin{table}[H]
	\caption{Simulation model parameters. For a full graphical model representation of the simulation scheme, including how these parameters are related, see Figure \ref{fig:sim-design}.}
	\begin{tabularx}{\linewidth}{p{1.5cm}ll} \toprule 
		Variable & Expression & Definition \\ \midrule 
		$\sigma^2_z$ &  0.73 log copies$^{2}$/mL$^{2}$ & Total spVL variance \\ 
		$H^2_h$ & 0.25 & Host heritability of spVL \\
		$H^2_{\bar{t}}$ & varied & Viral heritability of spVL at $\bar{t}$ \\
		$\sigma^2_{h}$ & $\sigma_{h}^2 = 0.25*\sigma_z^2$ & Variance in host part of spVL \\
		$\sigma^2_{p}(\bar{t})$ & $\sigma^2_{g_{v}}(\bar{t}) = H^2_{\bar{t}}*\sigma_z^2$ & Variance in viral part of spVL at $\bar{t}$ \\ 
		$\sigma^2_{\epsilon}$ & $\sigma_{\epsilon}^2 = \sigma_z^2 - \sigma^2_{p} - \sigma^2_{h}$ & Variance in environmental part of spVL \\ 
		$\bar{t}$ & 0.14 substitutions site$^{-1}$ yr$^{-1}$ & Mean root-tip time in viral phylogeny \\ 
		$\bm{g}_v$ & $\bm{g}_v \sim Norm(\bm{\mu}_{OU}, \boldsymbol{\Sigma}_{OU})$ & Viral part of spVL for all individuals \\ 
		$\theta$ & 4.5 log copies/mL & Optimal spVL value \\
		$g_{0}$ & 4.5 log copies/mL & $g$ at the root of the phylogeny  \\ 
		$\alpha$ & varied & Selection strength of OU process \\ 
		$\sigma$ & $\sigma = \sqrt{\frac{-2\alpha H^2 \sigma^2_z (0.75 - H^2)}{(H^2 - 1)(1 - exp(-2\alpha\bar{t}))}}$ & Time-unit standard deviation of OU process \\ 
		$\Psi$ & branch lengths $\sim Exp(\bar{t})$ & Viral phylogeny \\ 
		$g_{h_i}$ & \makecell{$g_{h_i} = \delta \sum_{j = 1}^{j = M/2}{G_{ij}} - \\ \delta \sum_{j = M/2}^{j = M}{G_{ij}}$} & Host part of spVL for individual $i$ \\ 
		$G_{N \times M}$ & \makecell{$G_{ij} \sim Binom(2, p)\ \\ \forall i \in {1...N}, \forall j \in {1...M}$} & Host genotype matrix \\
		$p$ & $p = \frac{1}{2} - \sqrt{\frac{1}{4} - \frac{H_{h}^2\sigma^2_z}{2\delta^2M}}$ &  Host variant allele frequency \\ 
		$\delta$ & 0.2 & Host variant effect size \\ 
		$M$ & 20 & Number of causal host variants \\ 
		$\epsilon_i$ & $ \epsilon_i \sim Norm(0, \sigma^2_{\epsilon})$ & Environmental part of spVL for individual $i$ \\ 
		$N$ & 500 & Number of simulated samples \\ \bottomrule 
	\end{tabularx}
	\label{tab:sim-params}
\end{table}

\subsection*{Swiss HIV-1 data}

Human genotypes, viral load measurements, and HIV-1 $pol$ gene sequences from HIV-1 positive individuals were all collected in the context of other studies by the Swiss HIV Cohort Study (SHCS) (www.shcs.ch, \citet{Schoeni-Affolter2010, Scherrer2021CohortSHCS}). All participants were HIV-1–infected individuals 16 years or older and written informed consent was obtained from all cohort participants. The anonymized data were made available for this study after the study proposal was approved by the SHCS.

For phylogenetic inference, we retained sequences from 1,493 individuals with non-recombinant subtype B $pol$ gene sequences of at least 750 characters and paired RNA measurements allowing for calculation of spVL, as well as 5 randomly chosen subtype A sequences as an outgroup. We used MUSCLE version 3.8.31 \citep{edgar_muscle:_2004} to align the $pol$ sequences with --maxiters 3 and otherwise default settings. We trimmed the alignment to 1505 characters to standardize sequence lengths. We used IQ-TREE version 1.6.9 \citep{Nguyen2014} to construct an approximate maximum likelihood tree with -m GTR+F+R4 for a general time reversible substitution model with empirical base frequencies and four free substitution rate categories. Otherwise, we used the default IQ-TREE settings. After rooting the tree based on the subtype A samples, we removed the outgroup. Viral subtype was determined by the SHCS using the REGA HIV subtyping tool version 2.0 \citep{DeOliveira2005}. We calculated spVL as the arithmetic mean of viral RNA measurements made prior to the start of antiretroviral treatment. For a comparison of several different filtering methods, see Figure \ref{fig:spvl-calc-comparison}.

For GWAS, we retained data from 1,392 of the 1,493 SHCS individuals with European ancestry who were not closely related to other individuals in the cohort (Table \ref{tab:sample-filtering}). These were 227 females and 1165 males. Ancestry was determined by plotting individuals along the three primary axes of genotypic variation from a combined dataset of SHCS samples and HapMap populations (Figure \ref{fig:PCA}). Kinship was evaluated using PLINK version 2.3 \citep{Chang2015}; we used the --king-cutoff option to exclude one from each pair of individuals with a kinship coefficient $>$ 0.09375. Initial host genotyping quality control and imputation were done as in \cite{Thorball2021GeneticLymphoma}. Subsequent genotyping quality control was performed using PLINK version 2.3 \citep{Chang2015}. We used the options --maf 0.01, --geno 0.01, and --hwe 0.00005 to remove variants with minor allele frequency less than 0.01, missing call rate greater than 0.05, or Hardy-Weinberg equilibrium exact test p-value less than 5x10$^-5$. After quality filtering, approximately 6.2 million genetic variants from the 1,392 individuals were retained for GWAS (Table \ref{tab:variant-filtering}).

\subsection*{\emph{A. thaliana}-\emph{X. arboricola} data}
\emph{A. thaliana} and \emph{X. arboricola} genotyping and quantitative disease resistance (QDR) measurements are described in detail in \citep{Wang2018Two-wayGenomes}. Briefly, \citep{Wang2018Two-wayGenomes} infected different \emph{A. thaliana} host accessions with different \emph{X. arboricola} pathogen strains in a fully-crossed experimental design. They infected up to 4 leaves on each of three biological replicates for each host-pathogen pair. Then, they scored QDR for each leaf on a scale of 0 (resistant) to 4 (susceptible). A genotype matrix with allele dosage of 33,610 SNPs for the 22 \emph{X. arboricola} pathogen strains used was downloaded from the supplemental material of \citep{Wang2018Two-wayGenomes}. A VCF file with allele dosage of 12,883,854 SNPs for different \emph{A. thaliana} accessions was downloaded from the 1001 Genomes project \citep{Alonso-Blanco20161135Thaliana}. QDR measurements were provided directly by the authors of \citep{Wang2018Two-wayGenomes}.

For phylogenetic inference, we used the ``dist.gene'' and ``nj'' functions from the ape package in R to construct a pairwise genetic distance matrix and then a neighbor-joining tree from the \emph{X. arboricola} pathogen genotype matrix. For the trait value to fit the POUMM, we calculated mean QDR across all leaves infected on all hosts for each \emph{X. arboricola} strain.

For GWAS, we retained genotype data from 132 \emph{A. thaliana} accessions with associated QDR measurements. We used the options --maf 0.1 and --max-alleles 2 to remove variants with minor allele frequency less than 0.1 as in \citep{Wang2018Two-wayGenomes} and multi-allelic variants. After filtering, approximately 1.7 million genetic variants from the 132 accessions were retained for GWAS (Table \ref{tab:variant-filtering-plants}). For the GWAS trait value, we randomly selected one pathogen pairing per host accession (weighted by number of leaves scored across the three replicates), so that the data is paired rather than fully-crossed and thus analogous to the HIV spVL GWAS analysis.

\subsection*{POUMM parameter inference}

We used the R package POUMM version 2.1.6 \citep{Mitov2017a-POUMM} to infer the POUMM parameters $g_0, \alpha, \theta, \sigma, \text{ and }\sigma_e$ from the HIV-1 and \emph{X. arboricola} phylogenies and associated spVL and QDR trait values. The Bayesian inference method implemented in this package requires specification of a prior distribution for each parameter. For HIV-1 spVL, we used the same, broad prior distributions as in \citet{Mitov2018}, namely: $g_{0} \sim \mathcal{N}(4.5,\ 3)$, $\alpha \sim Exp(0.02)$, $\theta \sim \mathcal{N}(4.5,\ 3)$, $H^2_{\bar{t}} \sim \mathcal{U}(0,\ 1)$, and $\sigma^2_e \sim Exp(0.02)$. For \emph{X. arboricola} QDR, we modified the $g_0$ and $\theta$ priors to match the empirical mean and standard deviation of QDR trait values in the dataset: $g_{0} \sim \mathcal{N}(0.4,\ 0.2)$ and $\theta \sim \mathcal{N}(0.4,\ 0.2)$. We ran two MCMC chains for 4x10$^6$ samples each with a target sample acceptance rate of 0.01 and a thinning interval of 1000 for both analyses. The first 2x10$^5$ samples of each chain were used for automatic adjustment of the MCMC proposal distribution. Figure \ref{fig:poumm-parameters} shows the posterior distributions for inferred parameters. Table \ref{tab:POUMM-parameters-our-estimates} gives the posterior mean values used for subsequent calculations.

\subsection*{Phylogenetic trait correction}
We corrected HIV-1 spVL and \emph{X. arboricola} QDR trait values using the method described in this paper. For each host individual, we estimated the pathogen part of their trait value using equation \ref{eq:MuG} and the corresponding non-pathogen part using equation \ref{eq:EHat}. For the POUMM parameters $\alpha$, $\sigma$, $\theta$, and $\sigma_e$, we used the posterior mean estimates generated as described above. As mentioned above, we randomly selected one \emph{X. arboricola} pathogen strain pairing per \emph{A. thaliana} host accession to be analogous to the human infectious disease case. Thus, trait values for GWAS were calculated as average QDR across three replicates for a particular pathogen-host pairing, while the strain-specific trait correction was calculated based on average QDR across all host pairings. All the code used to implement the method is available at https://github.com/cevo-public/POUMM-GWAS.

\subsection*{Association testing}
We performed two comparative GWAS for each infectious disease trait (HIV-1 spVL and \emph{X. arboricola} QDR). For the first ``GWAS with standard trait value'' we used the total (uncorrected) trait values ($\bm{z}$) as the response variable for association testing, replicating a standard GWAS set-up. For the second ``GWAS with estimated non-pathogen part of trait'' we replaced total trait values with the estimated non-pathogen component of the trait ($\hat{\bm{e}}$) as the response variable. Association testing was performed using a linear association model in PLINK version 2.3 \citep{Chang2015} with the top 5 principle components of host genetic variation included as covariates. In the HIV-1 spVL case, sex was also used as a covariate. These covariates were included to reduce residual variance in the trait values (sex) and control for confounding from host population structure (principle components).

\subsection*{Data availability}
The simulated data underlying this article can be re-generated using the code available on the project GitHub at https://github.com/cevo-public/POUMM-GWAS. The HIV pathogen genome sequences, clinical data, and human genotypes cannot be shared publicly due to the privacy of individuals who participated in the cohort study. The data may be shared on reasonable request to the Swiss HIV Cohort Study at http://www.shcs.ch. The \emph{X. arboricola} pathogen genotypes are available in the supplemental material of \citep{Wang2018Two-wayGenomes}, the \emph{A. thaliana} host genotypes are available at https://1001genomes.org/, and the \emph{A. thaliana}-\emph{X. arboricola} QDR measurements are available on request to the authors of \citep{Wang2018Two-wayGenomes}.

\section*{Acknowledgments}

This work was supported by ETH Zurich. We thank the patients who participate in the SHCS; the physicians and study nurses for excellent patient care; A. Scherrer,  E. Mauro, and K. Kusejko from the SHCS Data Centre for data management; and D. Perraudin and M. Amstad for administrative assistance. We thank Joy Bergelson for sharing the \emph{A. thaliana}-\emph{X. arboricola} QDR measurements. We also thank Michael Landis, who shared a LaTeX template for graphical model drawing.

The members of the SHCS are:
Abela I, Aebi-Popp K, Anagnostopoulos A, Battegay M, Bernasconi E, Braun DL, Bucher HC, Calmy A, Cavassini M, Ciuffi A, Dollenmaier G, Egger M, Elzi L, Fehr J, Fellay J, Furrer H, Fux CA, Günthard HF (President of the SHCS), Hachfeld A, Haerry D (deputy of "Positive Council"), Hasse B, Hirsch HH, Hoffmann M, Hösli I, Huber M, Kahlert CR (Chairman of the Mother & Child Substudy), Kaiser L, Keiser O, Klimkait T, Kouyos RD, Kovari H, Kusejko K (Head of Data Centre), Martinetti G, Martinez de Tejada B, Marzolini C, Metzner KJ, Müller N, Nemeth J, Nicca D, Paioni P, Pantaleo G, Perreau M, Rauch A (Chairman of the Scientific Board), Schmid P, Speck R, Stöckle M (Chairman of the Clinical and Laboratory Committee), Tarr P, Trkola A, Wandeler G, Yerly S.

The Swiss HIV Cohort Study is supported by the Swiss National Science Foundation (grant #201369), by SHCS project #858 and by the SHCS research foundation. Furthermore, the SHCS drug resistance database is supported by the Yvonne Jacob Foundation (to HFG). The data are gathered by the Five Swiss University Hospitals, two Cantonal Hospitals, 15 affiliated hospitals and 36 private physicians (listed in http://www.shcs.ch/180-health-care-providers).
% \end{doublespace}
\end{linenumbers}

\bibliographystyle{manuscript/my_natbib}
\bibliography{manuscript/references_synced}

%Additional ideas for supplemental materials:
%\begin{itemize}
%	\item Table of genotype samples by project (genotyping chip, number of samples, spVL distribution)
%	\item Link to github with all the code used to run simulations and generate figures
%	\item Example spVL plots
%	\item Proof of robustness of method to inferred tree -- NA because I just re-ran the _whole_ thing and it's not that different
%\end{itemize}

\newpage
\startsetups
    \setcounter{figure}{0}
    \renewcommand{\thefigure}{S\arabic{figure}}
    \setcounter{table}{0}
    \renewcommand{\thetable}{S\arabic{table}}
\stopsetups
\section*{Supplemental Material}\label{supp}
\beginsupplement
\setcounter{page}{1}

\begin{doublespace}

\subsection*{Expected results from simulations}

Here we show the root mean square error (RMSE) of $\bm{z} - \bm{\bar{z}}$ as an estimate for $\bm{h}$ should be $\approx$ 0.74 in our simulation scheme. First we write the expression for the RMSE:

\begin{equation}
    RMSE = \sqrt{\frac{\sum_i^N(z - \bar{z} - h)^2}{N}}
\end{equation}

Note that $z - \bar{z}$ differs from $h$ due to a viral effect and an environmental effect. So the term inside the square root equals the combined variance of these two effects:

\begin{equation}
    RMSE = \sqrt{\sigma^2_{p} + \sigma^2_\epsilon}
\end{equation}

We can calculate the variance due to these two effects because the total variance in spVL $\sigma^2_z$, and the fraction of the total variance due to host genetic effects, $\sigma^2_{h}$, are fixed parameters in our simulation scheme.

\begin{equation}
\begin{split}
    \sigma^2_{h} + \sigma^2_{p} + \sigma^2_\epsilon = \sigma^2_z \\
    0.25 * \sigma^2_{z} + \sigma^2_{p} + \sigma^2_\epsilon = \sigma^2_z \\
    \sigma^2_{p} + \sigma^2_\epsilon = 0.75 * \sigma^2_z \\
    \sigma^2_{p} + \sigma^2_\epsilon = 0.75 * 0.73 \\
    \sigma^2_{p} + \sigma^2_\epsilon = 0.55
\end{split}
\end{equation}

Therefore, we can expect the RMSE for $\bm{z} - \bm{\bar{z}}$ as an estimate for $\bm{h}$ to be around $\sqrt{0.55} \approx 0.74$.

\newpage

\begin{figure}[H]
	\begin{tikzpicture}

	% Variance 
	\begin{scope}[xshift = 0cm, yshift = 0cm]
	\node[anchor = west] (alabel) at (-1, 4) {\textbf{A} Variance partitioning};
	\node[constnode] (H2h) at (0, 0) {0.25};
	\node[constnode] (varz) at ($(H2h) + (1.8, 0)$) {$\sigma^2_z$};
	\node[constnode] (H2) at ($(H2h) + (3.6, 0)$) {\textcolor{red}{$H^2_{\bar{t}}$}};
	\node[detnode] (varh) at ($(H2h) + (0.9, 1.5)$) {$\sigma^2_{h}$};
	\node[detnode] (varv) at ($(varz) + (0.9, 1.5)$) {\tiny $\sigma^2_{g_{v}}(\bar{t})$};
	\node[detnode] (vare) at ($(varh) + (0.9, 1.5)$) {$\sigma^2_{\epsilon}$};
	\draw[taro] (H2h) -- (varh);
	\draw[taro] (varz) -- (varh);
	\draw[taro] (varz) -- (varv);
	\draw[taro] (H2) -- (varv);
	\draw[taro] (varh) -- (vare);
	\draw[taro] (varv) -- (vare);
	\draw[taro] (varz) -- (vare);
	\end{scope}

	% Virus
	\begin{scope}[xshift = 6cm, yshift = -0.5cm]
	\node[anchor = west] (blabel) at (-1, 4.5) {\textbf{B} Generating viral effects};
	\node[constnode] (theta) at (0, 1.5) {$\theta$};
	\node[constnode] (g0) at (1.8, 1.5) {$g_{0}$};
	\node[constnode] (alpha) at (3.6, 1.5) {\textcolor{red}{$\alpha$}};
	\node[detnode] (sigma) at (5.4, 1.5) {$\sigma$};
	\node[constnode] (tree) at (7.2, 1.5) {$\Psi$};
	\node[detnode] (varv) at (3.6+0.9, 0) {\tiny $\sigma^2_{g_{v}}(\bar{t})$};
	\node[constnode] (tbar) at (3.6+0.9+1.8, 0) {$\bar{t}$};
	\node[snode] (v) at (3.6, 3.25) {$\vec{p}$};
	\draw[taro] (theta) -- (v);
	\draw[taro] (g0) -- (v);
	\draw[taro] (alpha) -- (v);
	\draw[taro] (sigma) -- (v);
	\draw[taro] (tree) -- (v);
	\draw[taro] (varv) -- (sigma);
	\draw[taro] (tbar) -- (sigma);
	\draw[taro] (tbar) -- (tree);
	\draw[taro] (alpha) -- (sigma);
	\node[white] (varpic) at (3.6+0.9, -1.75) {};
	\draw[taro] (varpic) -- (varv);
	\node (vdist) at ($(v) + (1.8, 0.3)$) {Gaussian};
	\draw (v) -- (vdist);
	\draw[decoration={brace, mirror, raise=18pt}, sloped, decorate] (7.2, 1.5 - .5) -- node[below = 18pt] {OU parameters} (7.2, 1.5 + .5);
	\end{scope}

	% Host 
	\begin{scope}[xshift = 0cm, yshift = -7.5cm]
	\node[anchor = west] (clabel) at (-1, 5.55) {\textbf{C} Generating host effects};
	\node[constnode] (d) at (1.8, 1.5) {2};
	\node[detnode] (p) at (3.6, 1.5) {$g$};
	\node[snode] (G) at (1.8, 3) {$G$};
	\node[rectangle, dashed, very thick, inner sep=4mm, draw=black!100, fit = (G) ] (Gplate) {};
	\node[anchor=north west, inner sep=3pt] at (Gplate.north west) {\tiny $N\times M$};
	\node[detnode] (varh) at (3.6, 0) {$\sigma^2_{h}$};
	\node[constnode] (delta) at (3.6, 3) {$\delta$};
	\node[detnode] (h) at (3.6, 4.5) {$\vec{h}$};
	\node[constnode] (M) at (5.4, 3) {$M$};
	\draw[taro] (d) -- (G);
	\draw[taro] (p) -- (G);
	\draw[taro] (varh) -- (p);
	\draw[taro] (delta) -- (p);
	\draw[taro] (M) -- (p);
	\draw[taro] (M) -- (h);
	\draw[taro] (delta) -- (h);
	\draw[taro] (G) -- (h);
	\node[white] (varpic) at (3.6, -1.5) {};
	\draw[taro] (varpic) -- (varh);
	\node (hdist) at ($(h) + (2.7, 0.3)$) {$\approx$ Gaussian};
	\draw (h) -- (hdist);
	\node (gdist) at ($(G) + (-2, 0.6)$) {Binomial};
	\draw (G) -- (gdist);
	\end{scope}

	% Environment
	\begin{scope}[xshift = 9cm, yshift = -6.5cm]
	\node[anchor = west] (dlabel) at (-1, 3) {\textbf{D} Generating environmental effects};
	\node[detnode] (vare) at (0, 0) {$\sigma^2_{\epsilon}$};
	\node[constnode] (zero) at ($(vare) + (1.8, 0)$) {0};
	\node[snode] (e) at ($(vare) + (1.8/2, 1.5)$) {$\vec{\epsilon}$};
	\draw[taro] (zero) -- (e);
	\draw[taro] (vare) -- (e);
	\node[white] (varpic) at (0, -1.5) {};
	\draw[taro] (varpic) -- (vare);
	\node (edist) at ($(e) + (1.8, 0.3)$) {Gaussian};
	\draw (e) -- (edist);
	\end{scope}

	\end{tikzpicture}
	\caption{A graphical model representation of our simulation scheme, following the recommendations in \cite{Hohna2014}. Variables in solid squares are constants, with the two master control variables that we vary from simulation to simulation highlighted in red. Variables in solid circles are realizations of random variables and variables in dashed circles are determined as a function of other variables. Arrows represent dependencies among variables and the dashed square represents repetition. All parameters are defined in Table \ref{tab:sim-params}, as well as the values or expressions used for them. (A) shows how the variance in the simulated environmental effect $\sigma^2_{\epsilon}$ is smaller if the master pathogen heritability value $H^2_{\bar{t}}$ is higher and vice-versa. (B) shows the OU parameters and the pathogen phylogeny, which generate the Gaussian-distributed pathogen effects. The OU parameters $\theta$ and $g_0$ are fixed, whereas $\sigma$ is a deterministic function of the variance in the pathogen effect and the value of $\alpha$. In other words, we use $\sigma$ to maintain the desired pathogen heritability while varying $\alpha$. We generate a new random phylogeny for each simulation. (C) shows how host genotypes are drawn to generate host effects. The host genotype matrix $G$ contains the number of copies (0, 1, or 2) for each of $M$ causal variants with effect size $\delta$. We assume half the variants have a positive effect and half have a negative effect. The allele frequency $g$ for the causal variants set so that we achieve the desired variance in the host effects. (D) shows that the environmental effect is drawn from a Gaussian distribution with mean zero and variance as determined in part (A).}
	\label{fig:sim-design}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.7\linewidth]{figures/spvl_calculation_comparison.png}
	\caption{A comparison of different ways to calculate spVL based on viral load measurements provided by the SHCS. The stricter filtering excludes all measurements possibly $<$ 6 months after infection and after treatment or AIDS, whereas the more lenient filtering excludes only measurements after treatment. We used the lenient filter, mean measurement values because these correlate well with the values from the stricter filter but allow us to retain many more individuals from the cohort for our study.}
	\label{fig:spvl-calc-comparison}
	\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width = \linewidth]{figures/spvl_on_tree.png}
	\caption{Inferred HIV-1 $pol$ gene phylogeny with tips colored by (A) calculated spVL, (B) estimated non-pathogen effects on spVL and (C) estimated pathogen effects on spVL.}
	\label{fig:spvl-on-tree}
	\end{center}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width = 0.462\linewidth]{figures/trait_values.png}
	\caption{A comparison of measured (calculated) spVL values versus our estimated non-pathogen effect on spVL for each SHCS cohort member used in the study. The histograms show the marginal distribution of each value across the individuals.}
	\label{fig:trait-comparison}
	\end{center}
\end{figure}

\begin{figure}[H]
	\centering
		\includegraphics[width=0.7\linewidth]{figures/poumm_parameter_estimates.png}
		\caption{Posterior distributions compared to the prior for POUMM parameter estimates based on SHCS data. We ran two different MCMC chains to ensure the estimates converged. The black point on the x-axis shows the posterior mean value, which was used to estimate the pathogen- and non-pathogen effects on spVL.}
\label{fig:poumm-parameters}
\end{figure}

\begin{figure}[H]
	\centering
	\includegraphics[width=\linewidth]{figures/qq_plots.png}
	\caption{Quartile-quartile plots from association tests. The dashed green line shows the y = x line.}
	\label{fig:qq-plots}
\end{figure}

\begin{figure}[H]
	\includegraphics[width=\linewidth]{figures/host_genotype_pca.png}
	\caption{SHCS individuals and HapMap3 individuals plotted along the top three principle components of genetic variation. Points with black borders are within the thresholds used to select individuals of likely European ancestry.}
	\label{fig:PCA}
\end{figure}

\end{doublespace}  % don't doublespace tables

\newpage

\begin{table}[H]
\caption{POUMM parameter estimates for spVL based on SHCS data. HPD = Highest posterior density.}
	\begin{tabularx}{\linewidth}{XXX} \hline 
	Parameter & Posterior mean & 95\% HPD \\ \hline 
	${g_{0}}$ & 4.23 & (1.72, 6.71) \\
	$\theta$ & 4.47 & (4.37, 4.58) \\
	$\sigma$ & 5.25 & (2.37, 7.9) \\
	$\alpha$ & 57.65 & (19.49, 95.2) \\
	$\sigma_e$ & 0.54 & (0.43, 0.65) \\ 
	$H^2_{\bar{t}}$ & 0.45 & (0.24, 0.67) \\ \hline
	\end{tabularx}
	\label{tab:POUMM-parameters-our-estimates}
\end{table}

% latex table generated in R 4.1.1 by xtable 1.8-4 package
% Sat Oct 30 10:31:16 2021
\begin{table}[ht]
\caption{Effect size and p-values from the top most strongly associated variants in the $\emph{CCR5}$ and MCR regions from each of the two GWAS performed in our study. ``Standard'' means the GWAS with standard spVL trait values and ``Corrected'' means the GWAS with the estimated non-pathogen part of the trait. Entries above the dividing line are the top-associated variants from the ``Standard'' GWAS and entries below the dividing line are the top-associated variants from the ``Corrected'' GWAS. Many entries overlap between the two.}
\begin{tabularx}{\linewidth}{XXXXXXX}
  \hline
    Region & Position & Variant & Standard effect size & Standard p-value & Corrected effect size & Corrected p-value \\ 
  \hline
    $\emph{CCR5}$ & 46531144 & rs9845968 & -0.16 & $5.6 \times 10^{-9}$ & -0.083 & $1.2 \times 10^{-7}$ \\ 
    $\emph{CCR5}$ & 46537849 & rs867620 & -0.16 & $3.2 \times 10^{-9}$ & -0.085 & $6 \times 10^{-8}$ \\ 
    $\emph{CCR5}$ & 46539864 & rs11130092 & -0.16 & $1.1 \times 10^{-9}$ & -0.087 & $2.6 \times 10^{-8}$ \\ 
    $\emph{CCR5}$ & 46540932 & rs10865942 & -0.16 & $8.4 \times 10^{-9}$ & -0.081 & $4 \times 10^{-7}$ \\ 
    $\emph{CCR5}$ & 46541147 & rs7430431 & -0.17 & $9.2 \times 10^{-10}$ & -0.088 & $2.3 \times 10^{-8}$ \\ 
    MHC & 31274380 & rs9264942 & -0.21 & $4.5 \times 10^{-13}$ & -0.12 & $3.7 \times 10^{-13}$ \\ 
    MHC & 31321919 & rs1055821 & -0.33 & $9.4 \times 10^{-13}$ & -0.19 & $1.4 \times 10^{-12}$ \\ 
    MHC & 31380034 & rs112243036 & -0.32 & $9.9 \times 10^{-16}$ & -0.17 & $3.7 \times 10^{-14}$ \\ 
    MHC & 31391401 & rs4418214 & -0.34 & $2.4 \times 10^{-14}$ & -0.18 & $2.5 \times 10^{-12}$ \\ 
    MHC & 31400137 & rs138130755 & -0.46 & $1 \times 10^{-12}$ & -0.26 & $2.6 \times 10^{-12}$ \\ 
    MHC & 31400705 & rs138117378 & -0.46 & $1 \times 10^{-12}$ & -0.26 & $2.6 \times 10^{-12}$ \\ 
    MHC & 31402358 & rs148792134 & -0.46 & $1 \times 10^{-12}$ & -0.26 & $2.6 \times 10^{-12}$ \\ 
    MHC & 31409677 & rs140991764 & -0.46 & $1 \times 10^{-12}$ & -0.26 & $2.6 \times 10^{-12}$ \\ 
    \hline
    $\emph{CCR5}$ & 46531144 & rs9845968 & -0.16 & $5.6 \times 10^{-9}$ & -0.083 & $1.2 \times 10^{-7}$ \\ 
    $\emph{CCR5}$ & 46537849 & rs867620 & -0.16 & $3.2 \times 10^{-9}$ & -0.085 & $6 \times 10^{-8}$ \\ 
    $\emph{CCR5}$ & 46539864 & rs11130092 & -0.16 & $1.1 \times 10^{-9}$ & -0.087 & $2.6 \times 10^{-8}$ \\ 
    $\emph{CCR5}$ & 46541147 & rs7430431 & -0.17 & $9.2 \times 10^{-10}$ & -0.088 & $2.3 \times 10^{-8}$ \\ 
    $\emph{CCR5}$ & 46556835 & rs6808142 & 0.15 & $8.3 \times 10^{-8}$ & 0.082 & $3.3 \times 10^{-7}$ \\ 
    MHC & 31274380 & rs9264942 & -0.21 & $4.5 \times 10^{-13}$ & -0.12 & $3.7 \times 10^{-13}$ \\ 
    MHC & 31321919 & rs1055821 & -0.33 & $9.4 \times 10^{-13}$ & -0.19 & $1.4 \times 10^{-12}$ \\ 
    MHC & 31367874 & rs111281598 & -0.37 & $1.5 \times 10^{-12}$ & -0.22 & $2.1 \times 10^{-12}$ \\ 
    MHC & 31376266 & rs73400361 & -0.37 & $1.4 \times 10^{-12}$ & -0.22 & $2.1 \times 10^{-12}$ \\ 
    MHC & 31380034 & rs112243036 & -0.32 & $9.9 \times 10^{-16}$ & -0.17 & $3.7 \times 10^{-14}$ \\ 
   \hline
\end{tabularx}
\label{tab:gwas-top-associations}
\end{table}

\begin{table}[H]
\caption{Summary statistics for log spVL in previously sampled populations. $\bar{z}$ is average spVL (log copies/mL) and $\sigma^2_z$ is variance in measured spVL ( log $\textrm{copies}^2/\textrm{mL}^2$). Values from \protect\citep{Blanquart2017, Mitov2018} are empirical; values from \protect\citep{Bonhoeffer2015} were estimated by fitting a normal distribution to the data.}
	\begin{tabularx}{\linewidth}{XXX} \hline 
	Measurement & Value & Reference \\ \hline 
	$\bar{z}$ & $\approx$ 4.5 & \citet{Mitov2018} \\
	$\bar{z}$ &  4.4 & \citet{Blanquart2017} \\
	$\bar{z}$ & $\approx$ 4.5 & \citet{Bonhoeffer2015} \\
	$\sigma^2_z$ &  0.73 &  \citet{Mitov2018} \\
	$\sigma^2_z$ &  0.50 & \citet{Blanquart2017} \\
	$\sigma^2_z$ & $\approx$ 0.5 & \citet{Bonhoeffer2015} \\ \hline 
	\end{tabularx}
	\label{tab:spVLmeasurements}
\end{table}
%\todo{I guessed at bar(z) from figure in Mitov2018, search supplement for exact}

\begin{table}[H]
\caption{POUMM parameter estimates for spVL from previous studies.}
	\begin{tabularx}{\linewidth}{llXXX}
	\hline 
	Parameter & Value (Uncertainty) & Reference & Notes \\
	\hline 
	$g_0$ & 5.54 (4.04 - 7.25) & \citet{Mitov2018} & 8,483 UK HIV cohort individuals, $pol$ tree \\
	$\theta$ & 4.45 (4.41 - 4.49) & \citet{Mitov2018} & \\
	$\theta$ &  4.0 (1.6 - 4.)) & \citet{Bertels2018} & 3,036 SHCS individuals, $pol$ tree \\
	$\theta$ & 4.1 (3.5 - 4.9) & \citet{Blanquart2017} & 1,581 subtype B individuals from Europe, whole genome tree \\
	$\alpha$ & 28.78 (16.64 - 46.93) & \citet{Mitov2018} & \\
	$\alpha$ & 32.7 (0.03 - 57.6) & \citet{Bertels2018} & \\
	$\alpha$ & 7.6 (1.2 - 10) & \citet{Blanquart2017} & **limited $\alpha$ to $\le$ 10 \\
	$\sigma$ & 2.97 (1.95 - 4.37) & \citet{Mitov2018} & \\
	$\sigma$ & 1.3 (0.66 -1.87) &  \citet{Blanquart2017} & \\
	$\sigma_e$ & 0.77 (0.73, 0.8) & \citet{Mitov2018} & \\
	$\sigma_e$ & 0.61 (0.54, 0.65) &  \citet{Blanquart2017} & \\
	\hline 
	\end{tabularx}
	\label{tab:POUMMparams-prev-estimated}
\end{table}

\begin{table}[H]
\caption{Number of samples for GWAS after sequential filtering steps.}
    \begin{tabularx}{\linewidth}{XX}
    \hline
    Sample filter & Number of samples remaining \\
    \hline
    Subtype B pol sequences & 1516 \\
    With paired spVL measurement & 1516 \\
    $>$ 750 characters in sequence & 1493 \\
    Individual is of European ancestry & 1396 \\
    Kinship coefficient $>$ 0.09375 & 1392 \\ 
    \hline
    \end{tabularx}
    \label{tab:sample-filtering}
\end{table}

\begin{table}[H]
\caption{Number of variants for GWAS after sequential filtering steps.}
	\begin{tabularx}{\linewidth}{XX}
	\hline
	Variant filter & Number of variants remaining \\
	\hline 
	Raw data & 76979521 \\
	Missing genotype rate $>$ 0.05 & 11590002 \\
	Hardy-Weinburg exact test p-value $< 5 \times 10^{-5}$ & 11589246 \\
	Minor allele frequency $<$ 0.01 & 6228626 \\ \hline			
	\end{tabularx}
	\label{tab:variant-filtering}
\end{table}

\begin{table}[H]
\caption{Number of variants for \emph{X. arboricola} QDR GWAS after sequential filtering steps.}
	\begin{tabularx}{\linewidth}{XX}
	\hline
	Variant filter & Number of variants remaining \\
	\hline 
	Raw data & 12883854 \\
	Bi-allelic variants & 11769920 \\
	Minor allele frequency $<$ 0.1 & 1743952 \\ \hline
	\end{tabularx}
	\label{tab:variant-filtering-plants}
\end{table}

\stopsection
\end{document}